A Middleware for Symmetric Peer-to-Peer Applications

### 2023-03-09 tpds-23/tpds-23.tex
...

\documentclass[sn-mathphys,iicol]{sn-jnl}% Math and Physical Sciences Reference Style 
%\documentclass[default,iicol]{sn-jnl}% Default with double column layout

\jyear{2022}%

\raggedbottom
%%\unnumbered% uncomment this for unnumbered level heads

\usepackage{verbatim}
\usepackage{graphicx}

\usepackage{xspace}
\newcommand{\FC}       {Freechains\xspace}
\newcommand{\reps}     {\emph{reps}\xspace}
\newcommand{\onerep}   {\emph{1~rep}\xspace}
\newcommand{\nreps}[1] {\emph{#1~reps\xspace}}
\newcommand{\code}[1]  {\texttt{\small{#1}}}
\newcommand{\Xon} {$1{\rightarrow}N$\xspace}
\newcommand{\Xno} {$1{\leftarrow}N$\xspace}
\newcommand{\Xnn} {$N{\leftrightarrow}N$\xspace}
\newcommand{\Xoo} {$1{\leftrightarrow}1$\xspace}
\newcommand{\Xo}  {$1{\hookleftarrow}$\xspace}
\newcommand{\lin}[1]{(\emph{ln. #1}\xspace)}
\newcommand{\linx}[1]{\emph{ln. #1}\xspace}

@@

\title[Article Title]{
    Symmetric Peer-to-Peer Applications
}

\author{\fnm{Francisco} \sur{Sant'Anna} \email{francisco@ime.uerj.br}}
\affil{\orgdiv{Department of Computer Science}, \orgname{Rio de Janeiro State University}, \orgaddress{\country{Brazil}}}

\abstract{
In real-time networked applications, such as collaborative documents, users can
interact remotely and yet share the exact same experience as if they were using
a single machine.
%
In this work, we propose a middleware for \emph{symmetric peer-to-peer
applications} in which decentralized instances can broadcast asynchronous
events and yet conform to identical behavior.
%
Peers are allowed to join and leave the network at any time.
Application developers must adhere to a deterministic and stateless API.
%, and only supports pre-allocated memory.
%
The middleware is responsible for synchronizing the events in a global shared
timeline across the network.
Also, based on memory snapshots and deterministic simulation, a ``time machine''
can rollback conflicting peers to resynchronize their states.
%
We show that the middleware can handle applications with over 20 peers in a
heterogeneous topologies under high churns.
For instance, in a simulation scenario with a network latency of 50ms and 25
events per minute, the peers need to roll back and resynchronize only 3\% of
the time.
}

\keywords{
collaboration, determinism, peer-to-peer, time machine
}

%%\pacs[JEL Classification]{D8, H51}

%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}

@@

@@

@@

A common approach towards symmetric applications is to introduce a middleware
to orchestrate events and time in the network~\cite{gals,croquet}.
This way, applications developers can rely on middleware primitives to trigger
events, which are intercepted and synchronized in a global shared timeline
across the network.
Developers must also restrict themselves to deterministic and stateless calls
only, such that execution can be equally reproduced in all nodes according to
the shared timeline. %~\cite{TODO:det_reprod}.
However, current solutions depend on a central server to interconnect network
nodes and determine a shared timeline.

In this work, we propose \emph{symmetric peer-to-peer applications}, which
do not rely on central servers for coordination.
Peers in the application form a dynamic network graph and communicate only
with direct neighbours, as in typical unstructured peer-to-peer
networks~\cite{p2p.survey}.
Events are flooded in the graph and are applied with a small deadline delay to
compensate the network latency.
To deal with events received out of order or too late, a distributed time
machine can rollback peers to a previous state and reapply events in order and
in time.
Our main contribution is the design of a middleware to ensure that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.

We perform simulations with over 20 peers in a 5-hop mixed topology with cycles
and straight lines.
We vary the network latency and peer churn, as well as the frequency and
deadlines of events.
%
We then measure (a) the recurrence of time travels, (b) the real-time pace of
peers, and (c) the recoverability from churns, such that they validate our
goals above, respectively.
%
As results, we show that the last two goals are met even under extreme
conditions: peers exhibit a time mismatch of under $0.2\%$ and take less than
$1.5s$ to recover from churns.
%
For the first goal, we analyze the scenarios in which the middleware can meet
event deadlines with the desired performance.
As an example, peers need to roll back only $3\%$ of the time under a scenario
with a network latency of $50ms$ and 25 events per minute.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{algos}
  \caption{
    \label{fig.algos}
    Four approaches to symmetric distributed applications: Croquet, GALS,
    CRDTs, and this work.
    The vertical arrows in parallel represent the timelines in nodes in the
    network.
    The arrows crossing nodes represent communication between them.
  }
\end{figure*}

In this section, we revisit existing solutions for symmetric distributed
applications.
We focus on
    (a) how the network is organized,
    (b) how global time is determined, and
    (c) how events are propagated and applied.
Figure~\ref{fig.table} compares three selected works regarding these aspects.
%
At the end of this section, we also discuss related work on local and
distributed time machines.

@@

Many aspects may affect the performance of symmetric distributed applications,
such as
    the network latency,
    the number of nodes in the network,
    the rate of user events,
    and the application frame rate.
%
It is expected that centralized solutions exhibit a better performance in
comparison to peer-to-peer applications, since nodes are only 2 hops aways from
each other.
%
The performance of Croquet depends exclusively on the network latency, which
restricts the application to a maximum FPS.
For instance, since all frame ticks need to be transmitted over the network,
a latency of 50ms supports at most 20 FPS.
Also, the higher is the FPS, the higher is the network traffic and server load,
which may also degrade performance.
%
GALS detaches latency from FPS, since each node has its own timeline.
However, the protocol requires two roundtrips to the server, which degrades
performance considerably.
The authors claim to achieve a responsiveness of 350ms in applications with
25 nodes at 25 FPS, where 350ms represents the time it takes for all nodes to
apply an event since its initial occurrence.
%
As we discuss in Section~\ref{sec.eval}, the performance of our midlleware is
measured in terms of rollbacks, i.e., the more the peers need to resynchronize
and travel back in time, the lower is the performance of the application.
In a similar scenario with 21 peers at 50 FPS, and average of 5 hops with 50ms
latency, we show 3.1\% of rollbacks with responsiveness of 500ms, which we
classify as moderate performance (Figure~\ref{fig.baks}, framed rectangle in
the center).
As expected, the peer-to-peer performance is lower in comparison to centralized
solutions, but still viable in many scenarios, as detalied in
Section~\ref{sec.eval}.
%
Regarding CRDTs, recall that they only provide eventual consistency with no
notion of a shared timeline in which peers go through identical steps.
Therefore, measuring the performance of event responsiveness, maximum FPS, or
clock synchronization is meaningless, given that events can be applied locally
and broadcasted using a best-effort policy.

@@

In Figure~\ref{fig.middleware}, a peer is depicted as an application that uses
an API to communicate with the middleware.
The application itself is not aware of the network and has no direct access to
other peers, which communicate transparently through the middleware.
The application generates asynchronous events that go through the middleware,
which synchronizes them with a timestamp with a deadline delay ($\Delta$)
scheduled to the future, such that all peers can satisfy.
The middleware controls the execution of the application by issuing ticks and
synchronized events.
Note that ticks need not to be broadcast, only $\Delta$ events.

@@

The middleware assumes that each peer assigns itself a unique identifier (e.g.,
a combination its MAC \& IP addresses).
%, such as TODO~\cite{TODO:hash}.
The FPS must be the same in all peers to ensure bit-identical simulation.

Note that the use of callbacks is a common technique for event-driven
programming in Games, GUIs, and interactive applications in general, sharing
recognition with related techniques such as \emph{MVC},
\emph{Publish-Subscribe}, and \emph{Observer} patterns~\cite{meyer,nystrom}.
%
Furthermore, the default run-to-completion semantics of callbacks is key to
ensure the desired determinism, since unlike threads, they run sequentially and
atomically from start to end~\cite{events,threads}.

@@

\textbf{(b) Real-time pace of peers:}
All peers output their ticks continuously, allowing us to compare them in the
shared timeline.
Every time we see a greater tick never seen in the timeline before, we start to
count smaller ticks from other peers, which constitute time violations.
As an example, if all peers sum $1000$ of such smaller ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
We also start each peer with a random delay of up to $10s$ to force a
synchronization mismatch at the beginning.
%
Considering all scenarios, the average is negligibly under $0.2\%$, with no
significant variance.
Therefore, we omit the resulting table (akin to Figure~\ref{fig.baks}), which
would not provide further insights.

@@

$ vi test.c: #define TEST_OFFLINE
$ sudo ./test2-ana.sh 2>&1 | tee all2.log
$ lua5.3 test2-sheet.lua
\end{comment}

Our results show that goals (b) and (c) are met in all scenarios consistently,
and we consider them nearly optimal:
Regardless of the network latency, event rate, and delay multiplier,
    the variation in the pace of peers is under $0.2\%$ (goal b), while
    recovering peers take $1.3$ seconds to catch up in time (goal c).
As discussed in Section~\ref{sec.tml}.b, the middleware may take up to $1$
second to catch up, resulting in no more than 2 time travels to recover peers.

@@

@@

The middleware can handle applications with over 20 peers in a heterogeneous
topologies and under high churns.
%
We show that peers can
    (a) broadcast events and meet deadlines globally, 
    (b) advance in sync in real time, and
    (c) leave and join the network and remain symmetric.
%
In our experiments, we vary the network latency and the rate and deadline of
events.
We than measure the recurrence of time travels in the network as a quality
proxy.
%
As a result, we present a performance table to indicate the feasibility of
diverse application scenarios.
For instance, in a simulation scenario with a network latency of 50ms and 25
events per minute, the peers need to rollback only 3\% of the time.

\section*{Declarations}

\subsection*{Ethics Approval}

Not applicable

\subsection*{Conflict of Interest}

The author declares no conflicts of interest.

\subsection*{Data Availability}

Not applicable

\subsection*{Author Contribution}

Francisco Sant'Anna is the single author and wrote the full manuscript.

\subsection*{Funding}

This study was partially funded by FAPERJ (grant E-26/211.329/2019 (251640)).

\subsection*{Consent to publish}

The author hereby expresses consent to publish.

\bibliography{p2p-22}

>>>>

%\documentclass[journal]{IEEEtran}
\documentclass[10pt,journal,compsoc]{IEEEtran}

\usepackage{verbatim}
\usepackage{url}
\usepackage{graphicx}

\usepackage{xspace}
\newcommand{\reps}     {\emph{reps}\xspace}
\newcommand{\onerep}   {\emph{1~rep}\xspace}
\newcommand{\nreps}[1] {\emph{#1~reps\xspace}}
\newcommand{\code}[1]  {\texttt{\small{#1}}}
\newcommand{\Xon} {$1{\rightarrow}N$\xspace}
\newcommand{\Xno} {$1{\leftarrow}N$\xspace}
\newcommand{\Xnn} {$N{\leftrightarrow}N$\xspace}
\newcommand{\Xoo} {$1{\leftrightarrow}1$\xspace}
\newcommand{\Xo}  {$1{\hookleftarrow}$\xspace}
\newcommand{\lin}[1]{(\emph{ln. #1}\xspace)}
\newcommand{\linx}[1]{\emph{ln. #1}\xspace}

@@

\title{
    A Middleware for Symmetric Peer-to-Peer Applications
}

\author{
    Francisco Sant'Anna~\IEEEmembership{Department of Computer Science, Rio de Janeiro State University}
}

\IEEEtitleabstractindextext{%
\begin{abstract}
In real-time networked applications, such as collaborative documents, users can
interact remotely and yet share the exact same experience as if they were using
a single machine.
%
In this work, we propose a middleware for these \emph{symmetric peer-to-peer
applications} in which decentralized instances can broadcast asynchronous
events and yet conform to identical behavior.
%
Peers are allowed to join and leave the network at any time.
Application developers must adhere to a deterministic and stateless API.
%, and only supports pre-allocated memory.
%
The middleware is responsible for synchronizing the events in a global shared
timeline across the network.
Also, based on memory snapshots and deterministic simulation, a ``time machine''
can rollback conflicting peers to resynchronize their states.
%
We show that the middleware can handle applications with over 20 peers in
heterogeneous topologies under high churns.
For instance, in a scenario with a network latency of 50ms and 25 events per
minute, the peers need to roll back only 3\% of the time.
\end{abstract}

\begin{IEEEkeywords}
Bitcoin, blockchains, CRDT, distributed consensus, peer-to-peer,
publish-subscribe, reputation system, VCS
\end{IEEEkeywords}}

@@

% TOTAL: 12 pages

@@

 6. ***How does the middleware handle consistency when advancing or rollbacking in  time to synchronize?***
About #6, this is a valid criticism and we should include a brief discussion about consistency models and position our work on this spectrum.
The facts that peers share a global clock and go through bit-identical steps may characterize strict consistency.
However, since the global clocks may diverge temporarily, we must rely on rollbacks.
The key to preserve consistency in the presence of rollbacks is to restrict our programming API to deterministic and stateless calls only.
This way, it is not possible to diverge from other peers, for example, by re-executing stateful file operations multiple times (or at different times).

&gt; A apresentação da solução entra em um nível de detalhamento muito baixo...
  ...na decisão da topologia utilizada na simulação não fica muito claro se  essa topologia representa algum cenário específico.

Sobre a topologia escolhida, optamos por misturar 3 arranjos abstratos:

1. linha contínua
2. ciclo fracamente conectado
3. ciclo totalmente conectado

Entendemos que casos concretos que sejam subconjuntos dos arranjos acima terão desempenho igual ou superior ao exibido na Figura 7.
Por exemplo, uma linha contínua de até 15 nós é um subconjunto da topologia escolhida, assim como dois ciclos conectados separados por um nó.
Assim, pretendemos atender a um número maior de cenários, mesmo que isso dependa de uma interpretação mais apurada por parte dos interessados.
De qualquer maneira, a escolha é arbitrária e não representa um cenário concreto.

&gt; Figures 1 and 2 create an expectation of approaches comparison; I suggest writing in the motivation why there will be no clear comparison in the
  experimental results.
We will be more explicit about this in Sections 2/4, since it is not possible to reproduce our experiment in centralized solutions (the topology is inherently peer-to-peer).
We only discuss the performance of related works based on their claims and also on theoretical upper limits."

Os autores mencionam que em cenário com latência de 50ms e 25 eventos por minuto, os pares precisaram retroceder 3\% do tempo. Eles devem deixar claro porque isso é uma vantagem e qual a média de período de tempo aceitável para retrocesso.
- Nos trabalhos relacionados, apenas propostas centralizadas são mencionadas e comparadas. Não há propostas descentralizadas para comparação também? Sugiro expandir essa seção.
- O design experimental precisa ser melhor descrito. Apenas o emulador das latências é descrito (NetEm). Como foi realizada (ou qual ferramenta foi usada) a simulação dos experimentos?
- Apenas aplicações com deadline mais curtos adequam-se melhor ao middleware.  Os autores precisam explicar no texto os resultados dos experimentos com maior latência também (100ms, 250ms e 500ms).
- Os autores precisam deixar claro o benefício de usar uma solução descentralizada desde que o seu desempenho é menor que as soluções centralizadas.

@@

A common approach towards symmetric applications is to introduce a middleware
to orchestrate events and time in the network~\cite{gals,croquet}.
This way, applications developers can rely on middleware primitives to trigger
events, which are intercepted and synchronized in a global shared timeline
across the network.
Developers must also restrict themselves to deterministic and stateless calls
only, such that execution can be equally reproduced in all nodes according to
the shared timeline. %~\cite{TODO:det_reprod}.
However, current solutions depend on a central server to interconnect network
nodes and determine a shared timeline.
Therefore, applications may experience failures, unless users are permanently
connected, with no partitions in the network, particularly from the server.
In contrast, a decentralized solution can be resilient to partitions, and
ultimately, even work completely offline, as required by local-first
software~\cite{local}.

In this work, we propose a middleware for \emph{symmetric peer-to-peer
applications}, which do not rely on central servers for coordination.
Peers in the application form a dynamic network graph and communicate only
with direct neighbours, as in typical unstructured peer-to-peer
networks~\cite{p2p.survey}.
Events are flooded in the graph and are applied with a small deadline delay to
compensate the network latency.
To deal with events received out of order or too late, a distributed time
machine can rollback peers to a previous state and reapply events in order and
in time.
Our main contribution is the design of a middleware to ensure that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.

We perform simulations with over 20 peers in a 5-hop mixed topology with cycles
and straight lines.
We vary the network latency and peer churn, as well as the frequency and
deadlines of events.
%
We then measure (a) the recurrence of time travels, (b) the real-time pace of
peers, and (c) the recoverability from churns, such that they validate our
goals above, respectively.
%
As results, we show that goals (b) and (c) are met even under extreme
conditions: peers exhibit a time mismatch of under $0.2\%$ and take less than
$1.5s$ to recover from churns.
%
For the goal (a), we analyze the scenarios in which the middleware can meet
event deadlines with the desired performance.
As an example, peers need to roll back only $3\%$ of the time under a network
latency of $50ms$ and 25 events per minute.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{algos}
  \caption{
    \label{fig.algos}
    Four approaches to symmetric distributed applications: Croquet, GALS,
    CRDTs, and this work.
    The vertical arrows in parallel represent the nodes timelines.
    %in nodes in the network.
    The arrows crossing nodes represent communication between them.
  }
\end{figure*}

In this section, we revisit existing solutions for symmetric distributed
applications, which are not necessarily peer to peer.
We focus on
    (a) how the network is organized,
    (b) how global time is determined, and
    (c) how events are propagated and applied.
Figure~\ref{fig.table} compares three selected works regarding these aspects.
%
At the end of this section, we also discuss related work on local and
distributed time machines.

@@

@@

In Figure~\ref{fig.middleware}, a peer is depicted as an application using an
API to communicate with the middleware.
The application itself is not aware of the network and has no direct access to
other peers, which communicate transparently through the middleware.
The application generates asynchronous events that go through the middleware,
which synchronizes them with a timestamp with a deadline delay ($\Delta$)
scheduled to the future, such that all peers can satisfy.
The middleware controls the execution of the application by issuing ticks and
synchronized events.
Note that ticks need not to be broadcast, only $\Delta$ events.

@@

The middleware assumes that each peer assigns itself a unique identifier (e.g.,
a combination of its MAC \& IP addresses).
%, such as TODO~\cite{TODO:hash}.
The FPS must be the same in all peers to ensure bit-identical simulation.

Note that using callbacks is a common technique for event-driven
programming in Games, GUIs, and interactive applications in general, sharing
similarities with \emph{MVC}, \emph{Publish-Subscribe}, and \emph{Observer}
patterns~\cite{meyer,nystrom}.
%
Furthermore, the default run-to-completion semantics of callbacks is key to
ensure the desired determinism, since unlike threads, they run sequentially and
atomically from start to end~\cite{events,threads}.

@@

\textbf{(b) Real-time pace of peers:}
All peers output their ticks continuously, which we compare in the shared
timeline.
Every time we see a greater tick never seen in the timeline before, we start to
count smaller ticks from other peers, which constitute time violations.
As an example, if all peers sum $1000$ of such smaller ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
We also start each peer with a random delay of up to $10s$ to force a
synchronization mismatch at the beginning.
%
Considering all scenarios, the average is negligibly under $0.2\%$, with no
significant variance.
Therefore, we omit the resulting table (akin to Figure~\ref{fig.baks}), which
would not provide further insights.

@@

$ vi test.c: #define TEST_OFFLINE
$ sudo ./test2-ana.sh 2>&1 | tee all2.log
$ lua5.3 test2-sheet.lua
$
\end{comment}

Our results show that goals (b) and (c) are met in all scenarios consistently,
which are nearly optimal:
Regardless of the network latency, event rate, and delay multiplier,
    the variation in the pace of peers is under $0.2\%$ (goal b), while
    recovering peers take $1.3$ seconds to catch up in time (goal c).
As discussed in Section~\ref{sec.tml}.b, the middleware may take up to $1$
second to catch up, resulting in no more than 2 time travels to recover peers.

@@

\subsection{Discussion}

local-first
Apesar dos rollbacks quando a rede se reconecta serem inevitáveis, os seus efeitos podem ser imperceptíveis para os usuários finais (e.g., quando há pouca
atividade ou dependência entre eventos).

O principal benefício é que a aplicação continua funcionando mesmo com a
presença de partições na rede, o que inclui a queda do (que seria o) servidor
em uma solução centralizada.

Levando essa possibilidade ao extremo, um peer ainda poderia funcionar
normalmente mesmo que totalmente offline por um determinado período (como é
comum em pesquisas recentes em &quot;local-first software&quot;).

Many aspects may affect the performance of symmetric distributed applications,
such as
    the network latency,
    the number of nodes in the network,
    the rate of user events,
    and the application frame rate.
%
It is expected that centralized solutions exhibit a better performance in
comparison to peer-to-peer applications, since nodes are only 2 hops away from
each other.
%
However, the main benefit of our work is that applications remain functional
even under network partitions, which includes failures in (would be) servers.
Taking this possibility to an extreme, a peer can still work normally even
completely offline for a period of time, as desired by local-first
software~\cite{local}:
Even though rollbacks are inevitable when reconnecting, their effects may be
unnoticeable for final users if there is low activity or dependency between
events.
%
As pointed out in column (c) of Figure~\ref{fig.table}, centralized solutions
require that clients are permanently connected with the server, which controls
the shared timeline of events.

The performance of Croquet~\cite{croquet} depends exclusively on the network
latency, which restricts the application to a maximum FPS.
For instance, since all frame ticks need to be transmitted over the network,
a latency of 50ms supports at most 20 FPS.
Also, the higher is the FPS, the higher is the network traffic and server load,
which may also degrade performance.
%
GALS~\cite{gals} detaches latency from FPS, since each node has its own
timeline.
However, the protocol requires two round trips to the server, which degrades
performance considerably.
The authors claim to achieve a responsiveness of 350ms in applications with
25 nodes at 25 FPS, where 350ms represents the time it takes for all nodes to
apply an event since its initial occurrence.

The performance of our middleware is measured in terms of rollbacks, i.e., the
more the peers need to resynchronize and travel back in time, the lower is the
performance of the application.
In a similar scenario with 21 peers at 50 FPS, and average of 5 hops with 50ms
latency, we show 3.1\% of rollbacks with responsiveness of 500ms, which we
classify as moderate performance (Figure~\ref{fig.baks}, framed rectangle in
the center).
As expected, the peer-to-peer performance is lower in comparison to centralized
solutions, but still viable in many scenarios.

Regarding CRDTs, recall that they only provide eventual consistency with no
notion of a shared timeline in which peers go through identical steps.
Therefore, measuring the performance of event responsiveness, maximum FPS, or
clock synchronization is meaningless, given that events can be applied locally
and broadcast using a best-effort policy.

@@

The middleware can handle applications with over 20 peers in heterogeneous
topologies and under high churns.
%
We show that peers can
    (a) broadcast events and meet deadlines globally, 
    (b) advance in sync in real time, and
    (c) leave and join the network and remain symmetric.
%
In our experiments, we vary the network latency and the rate and deadline of
events.
We than measure the recurrence of time travels in the network as a quality
proxy.
%
As a result, we present a performance table to indicate the feasibility of
diverse application scenarios.
For instance, in a simulation scenario with a network latency of 50ms and 25
events per minute, the peers need to rollback only 3\% of the time.

\bibliographystyle{IEEEtran}
\bibliography{tpds-23}

%\begin{comment}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{chico}}]{Francisco Sant'Anna}
received his PhD degree in Computer Science from PUC-Rio, Brazil in
2013.
In 2016, he joined the Faculty of Computer Science at the Rio de Janeiro State
University, Brazil.
His research interests include Programming Languages and Concurrent \&
Distributed Systems.
\end{IEEEbiography}
%\end{comment}
